---
title: "R NFL Scraper Notebook"
output: 
  html_notebook:
    theme: paper
    toc: yes
    toc_float: yes
---

# Load packages {.tabset}

## tidyverse
```{r}
if (!require(tidyverse)){
  install.packages("tidyverse", dependencies = TRUE)
  library(tidyverse)
}
```

## devtools

```{r}
if (!require(devtools)){
  install.packages("devtools", dependencies = TRUE)
  library(devtools)
}
```
## rpart.plot
```{r}
if (!require(rpart.plot)){
  install.packages("rpart.plot", dependencies = TRUE)
  library(rpart.plot)
}
```


## nflScrapr

```{r}
if (!require(nflscrapR)){
  devtools::install_github(repo = "maksimhorowitz/nflscrapR")
  library(nflscrapR)
}
```

## caret

```{r}
if (!require(caret)){
  install.packages("caret", dependencies = TRUE)
  library(caret)
}
```

# Load Data {.tabset}

## download data from NFL
```{r}
games2009 <- season_games(Season = 2009)
games2010 <- season_games(Season = 2010)
games2011 <- season_games(Season = 2011)
games2012 <- season_games(Season = 2012)
games2013 <- season_games(Season = 2013)
games2014 <- season_games(Season = 2014)
games2015 <- season_games(Season = 2015)
games2016 <- season_games(Season = 2016)
games2017 <- season_games(Season = 2017)
gamesdf <- bind_rows(games2009, 
                     games2010, 
                     games2011, 
                     games2012, 
                     games2013, 
                     games2014, 
                     games2015, 
                     games2016,
                     games2017)

playerstats.2009 <- season_player_game(2009)
playerstats.2010 <- season_player_game(2010)
playerstats.2011 <- season_player_game(2011)
playerstats.2012 <- season_player_game(2012)
playerstats.2013 <- season_player_game(2013)
playerstats.2014 <- season_player_game(2014)
playerstats.2015 <- season_player_game(2015)
playerstats.2016 <- season_player_game(2016)
playerstats.2017 <- season_player_game(2017)
psdf <- bind_rows(playerstats.2009, 
                  playerstats.2010, 
                  playerstats.2011, 
                  playerstats.2012, 
                  playerstats.2013, 
                  playerstats.2014,
                  playerstats.2015, 
                  playerstats.2016,
                  playerstats.2017)
```
```{r}
games2018 <- season_games(Season = 2018)
playerstats.2018 <- season_player_game(2018)
```
```{r}
psdf1718 <- bind_rows(playerstats.2017, playerstats.2018)
gamesdf1718 <- bind_rows(games2017, games2018)
```

# Tidy data {.tabset}

## home team and opponent stats for 2018
```{r}
(playerstats.2018 %>%
  select(game.id, Team, passyds, rushyds, sacks, defints, fumbslost) %>%
  group_by(game.id, Team) %>%
  summarize(passyds = sum(passyds),
            rushyds = sum(rushyds),
            sacks = sum(sacks),
            defints = sum(defints),
            fumbslost = sum(fumbslost)) %>%
  arrange(game.id) %>%
  mutate(opassyds = case_when(row_number() %% 2 == 1 ~ passyds[2],
                               row_number() %% 2 == 0 ~ passyds[1]),
         orushyds = case_when(row_number() %% 2 == 1 ~ rushyds[2],
                              row_number() %% 2 == 0 ~ rushyds[1]),
         osacks = case_when(row_number() %% 2 == 1 ~ sacks[2],
                            row_number() %% 2 == 0 ~ sacks[1]),
         odefints = case_when(row_number() %% 2 == 1 ~ defints[2],
                              row_number() %% 2 == 0 ~ defints[1]),
         ofumbslost = case_when(row_number() %% 2 == 1 ~ fumbslost[2],
                                row_number() %% 2 == 0 ~ fumbslost[1])) %>%
  group_by(Team) %>%
  summarize(hpassyds = mean(passyds),
            hopassyds = mean(opassyds),
            hrushyds = mean(rushyds),
            horushyds = mean(orushyds),
            hsacks = mean(sacks),
            hosacks = mean(osacks),
            hfumbslost = mean(fumbslost),
            hofumbslost = mean(ofumbslost),
            hint = mean(defints), 
            hoint = mean(odefints)) %>%
  mutate(home = Team) %>%
  filter(!is.na(Team)) -> hfactorstats)
```
## visitor team and opponent stats for 2018
```{r}
(playerstats.2018 %>%
  select(game.id, Team, passyds, rushyds, sacks, defints, fumbslost) %>%
  group_by(game.id, Team) %>%
  summarize(passyds = sum(passyds),
            rushyds = sum(rushyds),
            sacks = sum(sacks),
            defints = sum(defints),
            fumbslost = sum(fumbslost)) %>%
  arrange(game.id) %>%
  mutate(opassyds = case_when(row_number() %% 2 == 1 ~ passyds[2],
                               row_number() %% 2 == 0 ~ passyds[1]),
         orushyds = case_when(row_number() %% 2 == 1 ~ rushyds[2],
                              row_number() %% 2 == 0 ~ rushyds[1]),
         osacks = case_when(row_number() %% 2 == 1 ~ sacks[2],
                            row_number() %% 2 == 0 ~ sacks[1]),
         odefints = case_when(row_number() %% 2 == 1 ~ defints[2],
                              row_number() %% 2 == 0 ~ defints[1]),
         ofumbslost = case_when(row_number() %% 2 == 1 ~ fumbslost[2],
                                row_number() %% 2 == 0 ~ fumbslost[1])) %>%
  group_by(Team) %>%
  summarize(vpassyds = mean(passyds),
            vopassyds = mean(opassyds),
            vrushyds = mean(rushyds),
            vorushyds = mean(orushyds),
            vsacks = mean(sacks),
            vosacks = mean(osacks),
            vfumbslost = mean(fumbslost),
            vofumbslost = mean(ofumbslost),
            vint = mean(defints), 
            voint = mean(odefints)) %>%
  mutate(away = Team) %>%
  filter(!is.na(Team)) -> vfactorstats)
```
## home team and opponent stats for 2009-2017
```{r}
(psdf %>%
  select(game.id, Team, passyds, rushyds, sacks, defints, fumbslost) %>%
  group_by(game.id, Team) %>%
  summarize(passyds = sum(passyds),
            rushyds = sum(rushyds),
            sacks = sum(sacks),
            defints = sum(defints),
            fumbslost = sum(fumbslost)) %>%
  arrange(game.id) %>%
  mutate(opassyds = case_when(row_number() %% 2 == 1 ~ passyds[2],
                               row_number() %% 2 == 0 ~ passyds[1]),
         orushyds = case_when(row_number() %% 2 == 1 ~ rushyds[2],
                              row_number() %% 2 == 0 ~ rushyds[1]),
         osacks = case_when(row_number() %% 2 == 1 ~ sacks[2],
                            row_number() %% 2 == 0 ~ sacks[1]),
         odefints = case_when(row_number() %% 2 == 1 ~ defints[2],
                              row_number() %% 2 == 0 ~ defints[1]),
         ofumbslost = case_when(row_number() %% 2 == 1 ~ fumbslost[2],
                                row_number() %% 2 == 0 ~ fumbslost[1])) %>%
  group_by(Team) %>%
  summarize(hpassyds = mean(passyds),
            hopassyds = mean(opassyds),
            hrushyds = mean(rushyds),
            horushyds = mean(orushyds),
            hsacks = mean(sacks),
            hosacks = mean(osacks),
            hfumbslost = mean(fumbslost),
            hofumbslost = mean(ofumbslost),
            hint = mean(defints), 
            hoint = mean(odefints)) %>%
  mutate(home = Team) %>%
  filter(!is.na(Team)) -> hfactorstatsOLD)
```
## visitor team and opponent stats for 2009-2017
```{r}
(psdf %>%
  select(game.id, Team, passyds, rushyds, sacks, defints, fumbslost) %>%
  group_by(game.id, Team) %>%
  summarize(passyds = sum(passyds),
            rushyds = sum(rushyds),
            sacks = sum(sacks),
            defints = sum(defints),
            fumbslost = sum(fumbslost)) %>%
  arrange(game.id) %>%
  mutate(opassyds = case_when(row_number() %% 2 == 1 ~ passyds[2],
                               row_number() %% 2 == 0 ~ passyds[1]),
         orushyds = case_when(row_number() %% 2 == 1 ~ rushyds[2],
                              row_number() %% 2 == 0 ~ rushyds[1]),
         osacks = case_when(row_number() %% 2 == 1 ~ sacks[2],
                            row_number() %% 2 == 0 ~ sacks[1]),
         odefints = case_when(row_number() %% 2 == 1 ~ defints[2],
                              row_number() %% 2 == 0 ~ defints[1]),
         ofumbslost = case_when(row_number() %% 2 == 1 ~ fumbslost[2],
                                row_number() %% 2 == 0 ~ fumbslost[1])) %>%
  group_by(Team) %>%
  summarize(vpassyds = mean(passyds),
            vopassyds = mean(opassyds),
            vrushyds = mean(rushyds),
            vorushyds = mean(orushyds),
            vsacks = mean(sacks),
            vosacks = mean(osacks),
            vfumbslost = mean(fumbslost),
            vofumbslost = mean(ofumbslost),
            vint = mean(defints), 
            voint = mean(odefints)) %>%
  mutate(away = Team) %>%
  filter(!is.na(Team)) -> vfactorstatsOLD)
```
## home team and opponent stats for 2017-2018
```{r}
(psdf1718 %>%
  select(game.id, Team, passyds, rushyds, sacks, defints, fumbslost) %>%
  group_by(game.id, Team) %>%
  summarize(passyds = sum(passyds),
            rushyds = sum(rushyds),
            sacks = sum(sacks),
            defints = sum(defints),
            fumbslost = sum(fumbslost)) %>%
  arrange(game.id) %>%
  mutate(opassyds = case_when(row_number() %% 2 == 1 ~ passyds[2],
                               row_number() %% 2 == 0 ~ passyds[1]),
         orushyds = case_when(row_number() %% 2 == 1 ~ rushyds[2],
                              row_number() %% 2 == 0 ~ rushyds[1]),
         osacks = case_when(row_number() %% 2 == 1 ~ sacks[2],
                            row_number() %% 2 == 0 ~ sacks[1]),
         odefints = case_when(row_number() %% 2 == 1 ~ defints[2],
                              row_number() %% 2 == 0 ~ defints[1]),
         ofumbslost = case_when(row_number() %% 2 == 1 ~ fumbslost[2],
                                row_number() %% 2 == 0 ~ fumbslost[1])) %>%
  group_by(Team) %>%
  summarize(hpassyds = mean(passyds),
            hopassyds = mean(opassyds),
            hrushyds = mean(rushyds),
            horushyds = mean(orushyds),
            hsacks = mean(sacks),
            hosacks = mean(osacks),
            hfumbslost = mean(fumbslost),
            hofumbslost = mean(ofumbslost),
            hint = mean(defints), 
            hoint = mean(odefints)) %>%
  mutate(home = Team) %>%
  filter(!is.na(Team)) -> hfactorstatsRECENT)
```
## visitor team and opponent stats for 2017-2018
```{r}
(psdf1718 %>%
  select(game.id, Team, passyds, rushyds, sacks, defints, fumbslost) %>%
  group_by(game.id, Team) %>%
  summarize(passyds = sum(passyds),
            rushyds = sum(rushyds),
            sacks = sum(sacks),
            defints = sum(defints),
            fumbslost = sum(fumbslost)) %>%
  arrange(game.id) %>%
  mutate(opassyds = case_when(row_number() %% 2 == 1 ~ passyds[2],
                               row_number() %% 2 == 0 ~ passyds[1]),
         orushyds = case_when(row_number() %% 2 == 1 ~ rushyds[2],
                              row_number() %% 2 == 0 ~ rushyds[1]),
         osacks = case_when(row_number() %% 2 == 1 ~ sacks[2],
                            row_number() %% 2 == 0 ~ sacks[1]),
         odefints = case_when(row_number() %% 2 == 1 ~ defints[2],
                              row_number() %% 2 == 0 ~ defints[1]),
         ofumbslost = case_when(row_number() %% 2 == 1 ~ fumbslost[2],
                                row_number() %% 2 == 0 ~ fumbslost[1])) %>%
  group_by(Team) %>%
  summarize(vpassyds = mean(passyds),
            vopassyds = mean(opassyds),
            vrushyds = mean(rushyds),
            vorushyds = mean(orushyds),
            vsacks = mean(sacks),
            vosacks = mean(osacks),
            vfumbslost = mean(fumbslost),
            vofumbslost = mean(ofumbslost),
            vint = mean(defints), 
            voint = mean(odefints)) %>%
  mutate(away = Team) %>%
  filter(!is.na(Team)) -> vfactorstatsRECENT)
```
## combine data for 2018
```{r}
games2018 %>%
  mutate(winner = case_when(homescore >= awayscore ~ "home",
                            homescore < awayscore ~ "away")) %>%
  left_join(hfactorstats) %>%
  left_join(vfactorstats, by = "away") %>%
  select(-GameID, -date, -home, -away, -homescore, -awayscore, -Team.x, -Team.y) -> dataset
```
## combine data for 2009-2017
```{r}
gamesdf %>%
  mutate(winner = case_when(homescore >= awayscore ~ "home",
                            homescore < awayscore ~ "away")) %>%
  left_join(hfactorstatsOLD) %>%
  left_join(vfactorstatsOLD, by = "away") %>%
  select(-GameID, -date, -home, -away, -homescore, -awayscore, -Team.x, -Team.y) -> datasetOLD
```
## combine data for 2017-2018
```{r}
gamesdf1718 %>%
  mutate(winner = case_when(homescore >= awayscore ~ "home",
                            homescore < awayscore ~ "away")) %>%
  left_join(hfactorstatsRECENT) %>%
  left_join(vfactorstatsRECENT, by = "away") %>%
  select(-GameID, -date, -home, -away, -homescore, -awayscore, -Team.x, -Team.y) -> datasetRECENT
```

# Create validiation dataset {.tabset} 

##for 2018

```{r}
# create a list of 80% of the rows in the original dataset we can use for training
validation_index <- createDataPartition(dataset$winner, p=0.80, list=FALSE)
# select 20% of the data for validation
validation <- dataset[-validation_index,]
# use the remaining 80% of data to training and testing the models
dataset <- dataset[validation_index,]
```

## for 2009-2017
```{r}
# create a list of 80% of the rows in the original dataset we can use for training
validation_indexOLD <- createDataPartition(datasetOLD$winner, p=0.80, list=FALSE)
# select 20% of the data for validation
validationOLD <- datasetOLD[-validation_indexOLD,]
# use the remaining 80% of data to training and testing the models
datasetOLD <- datasetOLD[validation_indexOLD,]
```
## for 2017-2018
```{r}
# create a list of 80% of the rows in the original dataset we can use for training
validation_indexRECENT <- createDataPartition(datasetRECENT$winner, p=0.80, list=FALSE)
# select 20% of the data for validation
validationRECENT <- datasetRECENT[-validation_indexRECENT,]
# use the remaining 80% of data to training and testing the models
datasetRECENT <- datasetRECENT[validation_indexRECENT,]
```

# Test Harness
```{r}
# Run algorithms using 10-fold cross validation
control <- trainControl(method="cv", number=10)
metric <- "Accuracy"
```

# Build 5 models {.tabset} 
##for 2018
```{r}
# a) linear algorithms
set.seed(7)
fit.lda <- train(winner~., data=dataset, method="lda", metric=metric, trControl=control)
# b) nonlinear algorithms
# CART
set.seed(7)
fit.cart <- train(winner~., data=dataset, method="rpart", metric=metric, trControl=control)
# kNN
set.seed(7)
fit.knn <- train(winner~., data=dataset, method="knn", metric=metric, trControl=control)
# c) advanced algorithms
# SVM
set.seed(7)
fit.svm <- train(winner~., data=dataset, method="svmRadial", metric=metric, trControl=control)
# Random Forest
set.seed(7)
fit.rf <- train(winner~., data=dataset, method="rf", metric=metric, trControl=control)
```
##for 2009-2017
```{r}
# a) linear algorithms
set.seed(7)
OLDfit.lda <- train(winner~., data=datasetOLD, method="lda", metric=metric, trControl=control)
# b) nonlinear algorithms
# CART
set.seed(7)
OLDfit.cart <- train(winner~., data=datasetOLD, method="rpart", metric=metric, trControl=control)
# kNN
set.seed(7)
OLDfit.knn <- train(winner~., data=datasetOLD, method="knn", metric=metric, trControl=control)
# c) advanced algorithms
# SVM
set.seed(7)
OLDfit.svm <- train(winner~., data=datasetOLD, method="svmRadial", metric=metric, trControl=control)
# Random Forest
set.seed(7)
OLDfit.rf <- train(winner~., data=datasetOLD, method="rf", metric=metric, trControl=control)
```
##for 2017-2018
```{r}
# a) linear algorithms
set.seed(7)
RECENTfit.lda <- train(winner~., data=datasetRECENT, method="lda", metric=metric, trControl=control)
# b) nonlinear algorithms
# CART
set.seed(7)
RECENTfit.cart <- train(winner~., data=datasetRECENT, method="rpart", metric=metric, trControl=control)
# kNN
set.seed(7)
RECENTfit.knn <- train(winner~., data=datasetRECENT, method="knn", metric=metric, trControl=control)
# c) advanced algorithms
# SVM
set.seed(7)
RECENTfit.svm <- train(winner~., data=datasetRECENT, method="svmRadial", metric=metric, trControl=control)
# Random Forest
set.seed(7)
RECENTfit.rf <- train(winner~., data=datasetRECENT, method="rf", metric=metric, trControl=control)
```
# Select best model {.tabset}
## 2018
```{r}
# summarize accuracy of models
results <- resamples(list(lda=fit.lda, cart=fit.cart, knn=fit.knn, svm=fit.svm, rf=fit.rf))
summary(results)
```
## 2009-2017
```{r}
# summarize accuracy of models
results <- resamples(list(lda=OLDfit.lda, cart=OLDfit.cart, knn=OLDfit.knn, svm=OLDfit.svm, rf=OLDfit.rf))
summary(results)
```
## 2017-2018
```{r}
# summarize accuracy of models
results <- resamples(list(lda=RECENTfit.lda, cart=RECENTfit.cart, knn=RECENTfit.knn, svm=RECENTfit.svm, rf=RECENTfit.rf))
summary(results)
```
# Test predictions
## 2018
```{r}
# estimate skill of LDA on the validation dataset
predictions <- predict(fit.rf, validation)
confusionMatrix(predictions, as_factor(validation$winner))
```
## 2009-2017
```{r}
# estimate skill of LDA on the validation dataset
predictionsOLD <- predict(OLDfit.rf, validationOLD)
confusionMatrix(predictionsOLD, as_factor(validationOLD$winner))
```
## 2017-2018
```{r}
# estimate skill of LDA on the validation dataset
predictionsRECENT <- predict(RECENTfit.rf, validationRECENT)
confusionMatrix(predictionsRECENT, as_factor(validationRECENT$winner))
```
# Make new predictions
```{r}
tibble::tribble(
       ~venue,    ~g1,    ~g2,    ~g3,    ~g4,    ~g5,    ~g6,    ~g7,    ~g8,    ~g9,   ~g10,   ~g11,   ~g12,   ~g13,  ~g14,  ~g15,
  "road team", "PHI",  "TB", "PIT", "LAC", "SEA", "CHI",  "AZ", "IND", "CAR", "BUF", "LAR", "JAX", "BAL", "KC", "SF",
  "home team", "NYG", "ATL", "CIN", "CLE", "OAK", "MIA", "MIN", "NYJ", "WAS", "HOU", "DEN", "DAL", "TEN", "NE", "GB"
  ) %>% 
  gather(venue) %>%
  mutate(home_away = row_number()%%2) %>%
  mutate(home_away = case_when(home_away==1 ~ "away",
                               home_away==0 ~ "home")) %>%
  spread(home_away, value) %>%
  left_join(hfactorstats) %>%
  left_join(vfactorstats, by = "away") -> currentWeek
```

```{r}
currentWeek %>%
  filter(venue == "g1")->currentWeek1

predict(RECENTfit.lda, currentWeek1)
```

# Plot decision tree for CART {.tabset}
## 2018
```{r}
rpart.plot(fit.cart$finalModel)
rpart.rules(fit.cart$finalModel)
rpart.predict(fit.cart$finalModel, currentWeek)
```


## 2009-2017
```{r}
rpart.plot(OLDfit.cart$finalModel)
rpart.rules(OLDfit.cart$finalModel)
rpart.predict(OLDfit.cart$finalModel, currentWeek)
```

## 2017-2018
```{r}
rpart.plot(RECENTfit.cart$finalModel)
rpart.rules(RECENTfit.cart$finalModel)
rpart.predict(RECENTfit.cart$finalModel, currentWeek)
```





